Parkinson's Disease cluster analysis

### 2016-05-21 writeup/manuscript/manuscript.tex
First round of manuscript corrections

% Use times font
\usepackage{mathptmx}
\usepackage{lmodern}

@@

% Tables, Figures
\usepackage{booktabs}
\usepackage{threeparttable}
% For line-breaking cells
\newcommand{\spc}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\usepackage{float}
\usepackage{caption}
\setlength{\tabcolsep}{10pt}

@@

% DOCUMENT ====
\begin{document}
% FRONTMATTER ====
\begin{frontmatter}
\title{Parkinson's disease cluster analysis}
% \author[c3]{Pablo Martinez-Martin}
% \ead{pmartinez@isciii.es}
% \author[bc]{Jesse Mu}
% \ead{muj@bc.edu}
% \author[cig]{Concha Bielza}
% \ead{mcbielza@fi.upm.es}
% \author[cig]{Pedro Larra\~naga}
% \ead{plarranaga@fi.upm.es}
% \address[c3]{Area of Applied Epidemiology, National Centre of Epidemiology and CIBERNED, Carlos III Institute of Health, Madrid, Spain}
% \address[bc]{Department of Computer Science, Boston College, Chestnut Hill, Massachusetts, USA}
% \address[cig]{Computational Intelligence Group, Polytechnic University of Madrid, Madrid, Spain}

@@

Out of the 951 patients in the study, we used listwise deletion to exclude 47 patients due to
missing measurements in the variables of interest, resulting in 904 remaining patients. All
symptoms were scaled to $z$-scores ($\mu = 0$, $\sigma = 1$) before clustering, and unscaled
afterwards for interpretation. Analyses were conducted in R 3.2.4 (www.r-project.org).

@@

$k$-means was used for cluster analysis. We performed two analyses on the
patients in the dataset: the first on the nine aggregate nonmotor symptom
domains and the four motor symptoms (``domains clustering''), and the second on the 30 individual
nonmotor symptoms collected with the four motor symptoms (``symptoms clustering''). Complete-linkage
hierarchical clustering on the 30 nonmotor symptoms was also performed to observe the
categorization of symptoms.

@@

Various formal measures were used to determine the optimal number of
clusters for the dataset. First, the optimal $k$ according to the Gap Statistic and the
1-standard-error method\cite{tibshirani01gap} was $k = 4$ (Figure~\ref{fig:gap}). Second, although
the globally-optimal clusters reported from stability measures of the clValid R package
\cite{brock11clvalid} stability were 2, 6, and 8 (Figure~\ref{fig:stability}), there appeared to be
consensus that $k = 4$ is a locally-optimal number of clusters which supported the gap statistic
result.

Other cluster validation methods (SSE scree plot, minimum average silhouette width) generally
suggested 2 clusters. As a consequence, $k = 2, 3, 4$ was tried, where $k = 2, 3$ simply divided
the data uninformatively into groups with varying levels of overall PD severity. Thus $k = 4$ was
selected to offer a good blend of model fit, informativeness, and parsimony.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.pdf}
  \caption{Plot of gap statistic versus number of clusters with $k$-means on 100 bootstrapped
    samples. Error bars are standard error. Per the method described in Tibshirani\cite{tibshirani01gap}, 4 is the smallest $k$ such
    that $\text{Gap}(k) \geq \text{Gap}(k + 1) - \text{se}_{k + 1}$.}
  \label{fig:gap}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{stability-measures.pdf}
  \caption{Average proportion of non-overlap (APN), average distance (AD), average distance
    between means (ADM), and figure of merit (FOM) stability measures\cite{datta2006methods,
    yeung2001validating}. Statistics should be minimized; notice the local optima at $k = 4$ for
  APN, AD, and ADM.}
  \label{fig:stability}
\end{figure}

@@

For each symptom in both clusterings, we used one-way ANOVA and $\chi^2$ tests to test the equality
of symptom means and proportions across the clusters found, using Bonferroni correction for
multiple testing with corrected $p < 0.05$ considered significant. Differences among pairwise
clusters were tested post-hoc using Tukey's range test for continuous means, or pairwise $\chi^2$
tests with Bonferroni correction for proportions.

To compare the clusterings, we computed the adjusted rand index\cite{hubert1985} (ARI) and
$v$-measure\cite{rosenberg07vmeasure} to evaluate similarity between the two clusterings, and used
stacked barplots to visualize the alignment of the clusters.

Lastly, to explore the relationship between symptom severity and disease duration, we computed the
linear correlation for each symptom with PD duration and fitted smoothed loess curves to the
data both globally and for each cluster found in the previous analyses.

@@

$k$-means clustering on the nine nonmotor domains and the four motor symptoms found four clusters
(Table~\ref{tab:nmd}; boxplots in Figure~\ref{fig:box}). Cluster means for all
symptoms were found to be statistically different $(p < 0.05 / 13 \approx 0.003)$, with specific
pairwise differences noted in the table. Additionally, statistics for variables not used in the
clustering were explored similarly in Table~\ref{tab:nmd_extra}; cluster means for all variables
were found to be statistically different, except for PD onset.

@@

Cluster 2 ($n = 189$) patients were severely affected in nonmotor domains but mildly affected in
motor domains. This cluster had an incidence of motor symptoms relatively similar to the Cluster 1
(mild) subtype especially in tremor, but generally expressed significantly higher scores for
nonmotor symptoms than Clusters 1 and 3, especially in the sleep/fatigue, mood/cognition and
miscellaneous domains. This group also had a statistically higher percentage of females than
Cluster 3.
% But also high variability

@@

\begin{table*}[t]
  \centering
  \caption{Summaries of domains clustering. Unless otherwise specified, statistics are
    reported as mean (sd). \textbf{Bold} indicates the highest score for that symptom out of the
  four clusters.}
  \label{tab:nmd}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    $n$ & 406 (45\%) & 189 (21\%) & 221 (24\%) & 88 (10\%) \\
    \midrule
    \textbf{Nonmotor} (Domains) & & & & \\
    Cardiovascular (0--24) & 0.7 (1.6)\tnote{24} & 2.3 (2.9)\tnote{134} & 1.1 (2.0)\tnote{24} &
    \textbf{7.0} (6.0)\tnote{234} \\
    Sleep/fatigue (0--48) & 4.3 (4.8)\tnote{234} & 14.9 (8.2)\tnote{134} & 6.9 (6.3)\tnote{124}
    & \textbf{21.0} (10.0)\tnote{123} \\
    Mood/cognition (0--60) & 3.1 (4.5)\tnote{234} & 16.1 (14.0)\tnote{134} & 6.6
    (8.1)\tnote{124} & \textbf{23.4} (14.3)\tnote{123} \\
    Perception/hallucination (0--36) & 0.5 (1.7)\tnote{24} & 1.9 (3.5)\tnote{134} & 0.8
    (2.0)\tnote{24} & \textbf{8.6} (6.9)\tnote{123} \\
    Attention/memory (0--36) & 2.8 (4.2)\tnote{24} & 8.9 (8.1)\tnote{134} & 3.2 (4.4)\tnote{24} &
    \textbf{15.6} (11.1)\tnote{123} \\
    Gastrointestinal (0--36) & 2.8 (3.9)\tnote{234} & 8.8 (7.1)\tnote{134} & 4.1
    (4.7)\tnote{124} & \textbf{14.6} (9.6)\tnote{123} \\
    Urinary (0--36) & 4.5 (5.9)\tnote{234} & 12.4 (9.7)\tnote{134} & 6.1 (6.4)\tnote{124} &
    \textbf{20.3} (9.9)\tnote{123} \\
    Sexual function (0--24) & 1.5 (3.1)\tnote{24} & 6.4 (7.4)\tnote{134} & 2.5 (4.1)\tnote{24} &
    \textbf{9.0} (9.7)\tnote{123} \\
    Miscellaneous (0--48) & 3.9 (4.7)\tnote{24} & 13.2 (8.6)\tnote{13} & 5.2 (5.8)\tnote{24} &
    \textbf{14.0} (9.5)\tnote{13} \\
    \midrule
    \textbf{Motor} & & & & \\
    Tremor & 1.9 (1.8)\tnote{34} & 1.6 (1.9)\tnote{34} & \textbf{4.3} (2.8)\tnote{124} & 3.5
    (3.8)\tnote{123} \\
    Bradykinesia & 1.5 (0.9)\tnote{234} & 2.1 (1.1)\tnote{134} & 3.6 (0.9)\tnote{124} &
    \textbf{4.0} (1.5)\tnote{123} \\
    Rigidity & 1.5 (0.9)\tnote{234} & 1.8 (1.1)\tnote{134} & 3.3 (0.9)\tnote{124} &
    \textbf{3.8} (1.4)\tnote{123} \\
    Axial & 1.8 (1.6)\tnote{234} & 3.3 (2.1)\tnote{134} & 4.3 (2.4)\tnote{124} &
    \textbf{7.4} (2.9)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[b]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4-pub.pdf}
  \vspace{-7.5em}
  \captionsetup{justification=raggedleft,
    singlelinecheck=false
  }
  \caption{Boxplot cluster summaries \\ for domains clustering.}
  \label{fig:box}
\end{figure*}

\begin{table*}[t]
  \centering
  \caption{Cluster statistics for variables not used in the analysis}
  \label{tab:nmd_extra}
  \begin{threeparttable}
  \begin{tabular}{l r r r r}
    \toprule
    Cluster & 1 & 2 & 3 & 4 \\
    \midrule
      Sex (\% male) & 63 & 53\tnote{3} & 71\tnote{2} & 58 \\
      Age & 62.7 (9.7)\tnote{34} & 64 (9.0)\tnote{4} & 65 (10.2)\tnote{14} & 70.5 (8.9)\tnote{123} \\
      PD\_Onset & 56 (10.6) & 55.2 (10.1) & 56.9 (11.1) & 58.2 (11.4) \\
      PD\_Duration & 6.7 (4.8)\tnote{234} & 8.7 (5.7)\tnote{14} & 8 (5.5)\tnote{14} & 12.3
      (8.1)\tnote{123} \\
      CISI\_Total & 5.7 (3.2)\tnote{234} & 9.2 (3.9)\tnote{14} & 9.7 (3.5)\tnote{14} & 14.8
      (5.1)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (adjusted $p < 0.05$)
    \item[2] Significant difference with cluster 2 (adjusted $p < 0.05$)
    \item[3] Significant difference with cluster 3 (adjusted $p < 0.05$)
    \item[4] Significant difference with cluster 4 (adjusted $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

@@

$k$-means clustering performed on the 30 individual nonmotor symptoms found similar patterns to the
first clustering (Table~\ref{tab:nms30}, heatmap in Figure~\ref{fig:nms30-hm}). Means of all
symptoms were found to differ across clusters except for PD\_Onset and tremor.

Clusters 1 ($n = 509$) and 4 ($n = 49$), the mild and severe subtypes of the previous clustering,
retained their characteristics here, while clusters 2 ($n = 97$) and 3 ($n = 249$) were different.
2 was a mood/cognition-dominant cluster, with a relatively higher proportion of female patients and
additionally a more severe incidence of insomnia. 3 was not characterized by any especially high
symptoms and was thus classified as a group of patients with average severity.
% Once again, CISI TOTAL...

@@

The two clusterings are somewhat similar ($\text{ARI} = 0.46$; $\text{$v$-measure} = 0.43$).
Alignment of the clusters is visualized in Figure~\ref{fig:align}. While clusters 1 and 4 of both
clusterings align well with each other, main differences in the clusters occur in the mixing of
clusters 2 and 3 of both clusterings; for example, domains cluster 3 (motor-dominant) is somewhat
evenly split among symptoms clusters 1 (mild) and 3 (average).

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{cluster-alignment.pdf}
  \caption{Bar chart indicating the distribution of opposite cluster membership for the patients of
  a given cluster. For example, the bar furthest left indicates that close to 100\% of patients in
domains cluster 1 are in symptoms cluster 1, with the rest in symptoms cluster 3.}
  \label{fig:align}
\end{figure}

@@

    statistics are reported as mean (sd). \textbf{Bold} indicates the highest score for that
  symptom out of the four clusters.}

@@

  Lightheadedness & 0.6 (1.3)\tnote{234} & 2.4 (3.0)\tnote{14} & 2.1 (2.9)\tnote{14} & \textbf{5.2}

@@

  &Fainting & 0.1 (0.7)\tnote{24} & 0.5 (1.4)\tnote{134} & 0.2 (0.6)\tnote{24} & \textbf{2.4}

@@

  Drowsiness & 1.0 (1.9)\tnote{234} & 2.8 (3.3)\tnote{14} & 2.5 (3.0)\tnote{14} & \textbf{7.0}

@@

  &Fatigue & 1.3 (2.0)\tnote{234} & 6.3 (4.2)\tnote{13} & 4.8 (3.9)\tnote{124} & \textbf{7.3} (3.5)\tnote{13} \\
  &Insomnia & 1.2 (2.6)\tnote{234} & \textbf{5.5} (4.8)\tnote{13} & 3.2 (3.8)\tnote{12} & 4.2

@@

  &RLS & 0.7 (1.6)\tnote{234} & 2.3 (3.5)\tnote{14} & 1.9 (3.2)\tnote{14} & \textbf{3.9}

@@

  3.\ Mood/cognition &
  Loss\_interest & 0.4 (1.1)\tnote{234} & \textbf{6.1} (3.8)\tnote{134} & 1.0 (1.8)\tnote{124} & 4.6 (3.7)\tnote{123} \\
  &Loss\_activities & 0.7 (1.6)\tnote{234} & \textbf{7.1} (4.1)\tnote{134} & 1.7 (2.4)\tnote{124} &

@@

  &Anxiety & 0.9 (1.8)\tnote{234} & \textbf{6.6} (4.3)\tnote{134} & 2.0 (2.8)\tnote{124} & 4.6 (3.9)\tnote{123} \\
  &Depression & 0.8 (1.6)\tnote{234} & \textbf{7.8} (3.7)\tnote{134} & 2.3 (3.0)\tnote{124} & 4.3 (3.7)\tnote{123} \\
  &Flat\_affect & 0.4 (1.1)\tnote{234} & \textbf{4.9} (4.3)\tnote{134} & 0.9 (1.9)\tnote{124} & 2.7 (3.0)\tnote{123} \\
  &Loss\_pleasure & 0.4 (1.2)\tnote{234} & \textbf{6.8} (4.1)\tnote{134} & 1.1 (2.1)\tnote{124} & 3.4 (3.3)\tnote{123} \\

@@

  4.\ Perception/ & Hallucination & 0.2 (1.0)\tnote{234} & 0.8 (1.9)\tnote{14} & 0.6 (1.6)\tnote{14} & \textbf{4.6} (3.6)\tnote{123} \\

@@

  \textbf{3.1} (3.8)\tnote{123} \\
  &Diplopia & 0.2 (0.7)\tnote{234} & 1.0 (2.5)\tnote{14} & 0.6 (1.4)\tnote{14} & \textbf{4.4}

@@

  5.\ Attention/ & Loss\_concentration & 0.9 (1.8)\tnote{234} & 4.0 (3.6)\tnote{134} & 2.5 (3.1)\tnote{124} & \textbf{6.6} (3.9)\tnote{123} \\
  memory &Forget\_explicit & 0.9 (1.6)\tnote{234} & 3.8 (3.9)\tnote{134} & 2.2 (2.9)\tnote{124} & \textbf{6.7} (3.6)\tnote{123} \\
  &Forget\_implicit & 0.7 (1.6)\tnote{234} & 3.1 (3.7)\tnote{134} & 1.8 (2.6)\tnote{124} & \textbf{6.7} (4.0)\tnote{123} \\

@@

  Drooling & 0.7 (1.6)\tnote{234} & 3.0 (4.1)\tnote{14} & 3.0 (3.8)\tnote{14} & \textbf{6.1} (4.4)\tnote{123} \\
  &Swallowing & 0.4 (1.2)\tnote{234} & 1.2 (2.1)\tnote{14} & 1.7 (2.8)\tnote{14} & \textbf{4.5} (3.9)\tnote{123} \\
  &Constipation & 1.6 (2.9)\tnote{234} & 3.6 (4.3)\tnote{14} & 3.3 (4.1)\tnote{14} & \textbf{7.0} (5.1)\tnote{123} \\

@@

  Urinary\_urgency & 1.0 (1.8)\tnote{234} & 3.5 (4.1)\tnote{14} & 3.8 (3.8)\tnote{14} & \textbf{7.8} (4.0)\tnote{123} \\
  &Urinary\_frequency & 0.9 (1.9)\tnote{234} & 4.0 (4.1)\tnote{14} & 3.8 (4.0)\tnote{14} & \textbf{6.5} (4.0)\tnote{123} \\
  &Nocturia & 1.7 (2.3)\tnote{234} & 5.0 (4.3)\tnote{14} & 5.1 (4.1)\tnote{14} & \textbf{7.7} (3.9)\tnote{123} \\

@@

  Sex\_drive & 0.8 (1.9)\tnote{234} & 4.5 (4.9)\tnote{13} & 2.3 (3.6)\tnote{124} & \textbf{5.2} (5.4)\tnote{13} \\

@@

  \textbf{4.6} (5.4)\tnote{13} \\

@@

  Unexplained\_pain & 0.7 (1.9)\tnote{234} & 2.5 (3.7)\tnote{14} & 3.0 (4.1)\tnote{14} & \textbf{4.2} (4.3)\tnote{123} \\

@@

  \textbf{4.7} (4.8)\tnote{13} \\
  &Weight\_change & 0.9 (1.7)\tnote{234} & 2.5 (3.4)\tnote{14} & 1.9 (3.0)\tnote{14} & \textbf{3.9} (4.2)\tnote{123} \\
  &Sweating & 0.7 (1.8)\tnote{234} & 2.2 (3.5)\tnote{1} & 2.8 (4.1)\tnote{1} & \textbf{3.2}

@@

  Motor &
  Tremor & 2.5 (2.3) & 2.1 (2.4) & 2.6 (2.6) & \textbf{4.1} (4.3) \\
  &Bradykinesia & 2.0 (1.2)\tnote{234} & 2.8 (1.4)\tnote{14} & 2.8 (1.4)\tnote{14} & \textbf{3.9} (1.8)\tnote{123} \\

>>>>

% Use times font
\usepackage{mathptmx}
\usepackage{lmodern}
\usepackage{times}

@@

% Tables
\usepackage{booktabs}
\usepackage{threeparttable}
% For line-breaking cells
\newcommand{\spc}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
% Reduce table spacing
\setlength{\tabcolsep}{10pt}

% Figures
\usepackage{float}
% Reset numbering for supplementary figures
\usepackage{newfloat}
\DeclareFloatingEnvironment[name={Supplementary Figure}]{suppfigure}
\usepackage{caption}

@@

% DOCUMENT ====
\begin{document}
% FRONTMATTER ====
\begin{frontmatter}
\title{Parkinson's Disease cluster analysis}
% \author[c3]{Pablo Martinez-Martin}
% \ead{pmartinez@isciii.es}
% \author[bc]{Jesse Mu}
% \ead{muj@bc.edu}
% \author[cig]{Concha Bielza}
% \ead{mcbielza@fi.upm.es}
% \author[cig]{Pedro Larra\~naga}
% \ead{plarranaga@fi.upm.es}
% \address[c3]{Area of Applied Epidemiology, National Centre of Epidemiology and CIBERNED, Carlos III Institute of Health, Madrid, Spain}
% \address[bc]{Department of Computer Science, Boston College, Chestnut Hill, Massachusetts, USA}
% \address[cig]{Computational Intelligence Group, Polytechnic University of Madrid, Madrid, Spain}

@@

Out of the 951 patients in the study, we used listwise deletion to exclude 47 patients due to
missing measurements in the variables of interest, resulting in 904 remaining patients. All
symptoms were standardized before clustering, and unstandardized afterwards for interpretation.
Analyses were conducted in R version 3.2.4 (www.r-project.org).

@@

$k$-means was used for cluster analysis. We performed two analyses on the
patients in the dataset: the first ``domains clustering'' on the nine aggregate nonmotor symptom
domains and the four motor symptoms, and the second ``symptoms clustering'' on the 30 individual
nonmotor symptoms collected with the four motor symptoms. Complete-linkage agglomerative
hierarchical clustering on the 30 nonmotor symptoms was also performed to how the symptoms group
together.

@@

Various formal measures were used to determine the optimal number of clusters for the dataset. The
optimal $k$ according to the Gap Statistic and the 1-standard-error method suggested by
Tibshirani\cite{tibshirani01gap} was $k = 4$ (Supplementary Material Figure~\ref{fig:gap}). Other cluster validation
methods (within sum of squares (WSS) scree plot, minimum average silhouette width) suggested $k =
2, 3, 4$, where $k = 2, 3$ simply divided the data uninformatively into groups with varying levels
of overall PD severity. Thus $k = 4$ was selected to offer a good blend of model fit,
informativeness, and parsimony.

@@

For each symptom in both clusterings, we used one-way ANOVA and $\chi^2$ tests to respectively
check the equality of symptom means and proportions across the clusters found, using Bonferroni
correction for multiple testing with corrected $p < 0.05$ considered significant. Differences among
pairwise clusters were tested post-hoc using Tukey's range test for continuous means, or pairwise
$\chi^2$ tests with Bonferroni correction for proportions.

To compare the clusterings, we visualized cluster alignment with stacked barplots, and computed the adjusted rand index\cite{hubert1985} (ARI) and
Rosenberg's $v$-measure.\cite{rosenberg07vmeasure} to evaluate similarity between the two
clusterings. Both measures range from 0 (no similarity) to 1 (identical), and respectively take a
contingency table and information-theoretic approach to measuring clustering similarity.

Lastly, to explore the relationship between symptom severity and disease duration, we computed the
correlation coefficient for each symptom with PD duration and fitted smoothed loess curves to the
data both globally and for each cluster found in the previous analyses.

@@

Results from the clustering on the nine nonmotor domains and the four motor symptoms, as well as
additional variables not used in the analysis, are reported in Table~\ref{tab:nmd}, with boxplots
in Figure~\ref{fig:box}.
Cluster means for all variables were found to be statistically different, except for PD onset ($p <
0.05 / 18 \approx 0.003$ after correcting for the comparisons of the 18 variables). Specific
pairwise differences are noted in the table.

@@

Cluster 2 ($n = 189$) patients were severely affected in nonmotor domains but mildly affected in
motor domains. This cluster had an incidence of motor symptoms relatively similar to the Cluster 1
(mild) subtype especially in tremor, but generally expressed significantly higher scores for
nonmotor symptoms than Clusters 1 and 3, especially in the sleep/fatigue, mood/apathy and
miscellaneous domains. This group also had a statistically higher percentage of females than
Cluster 3.
% But also high variability

@@

\begin{table*}[t]
  \centering
  \caption{Domains clustering summary. Unless otherwise specified, statistics are
  reported as mean (sd).}
  \label{tab:nmd}
  \begin{threeparttable}
    \small
    \begin{tabular}{lrllll}
    \toprule
    & Cluster & 1 & 2 & 3 & 4 \\
    & $n$ & 406 (45\%) & 189 (21\%) & 221 (24\%) & 88 (10\%) \\
    \midrule
    Nonmotor domains & Cardiovascular & 0.7 (1.6)\tnote{24} & 2.3 (2.9)\tnote{134} & 1.1 (2.0)\tnote{24} &
    7.0 (6.0)\tnote{123} \\
    & Sleep/fatigue & 4.3 (4.8)\tnote{234} & 14.9 (8.2)\tnote{134} & 6.9 (6.3)\tnote{124}
    & 21.0 (10.0)\tnote{123} \\
  & Mood/apathy & 3.1 (4.5)\tnote{234} & 16.1 (14.0)\tnote{134} & 6.6
    (8.1)\tnote{124} & 23.4 (14.3)\tnote{123} \\
    & Perception/hallucination & 0.5 (1.7)\tnote{24} & 1.9 (3.5)\tnote{134} & 0.8
    (2.0)\tnote{24} & 8.6 (6.9)\tnote{123} \\
    & Attention/memory & 2.8 (4.2)\tnote{24} & 8.9 (8.1)\tnote{134} & 3.2 (4.4)\tnote{24} &
    15.6 (11.1)\tnote{123} \\
    & Gastrointestinal & 2.8 (3.9)\tnote{234} & 8.8 (7.1)\tnote{134} & 4.1
    (4.7)\tnote{124} & 14.6 (9.6)\tnote{123} \\
    & Urinary & 4.5 (5.9)\tnote{234} & 12.4 (9.7)\tnote{134} & 6.1 (6.4)\tnote{124} &
    20.3 (9.9)\tnote{123} \\
    & Sexual function & 1.5 (3.1)\tnote{24} & 6.4 (7.4)\tnote{134} & 2.5 (4.1)\tnote{24} &
    9.0 (9.7)\tnote{123} \\
    & Miscellaneous & 3.9 (4.7)\tnote{24} & 13.2 (8.6)\tnote{13} & 5.2 (5.8)\tnote{24} &
    14.0 (9.5)\tnote{13} \\
    \midrule
    Motor symptoms & Tremor & 1.9 (1.8)\tnote{34} & 1.6 (1.9)\tnote{34} & 4.3 (2.8)\tnote{124} & 3.5
    (3.8)\tnote{123} \\
    & Bradykinesia & 1.5 (0.9)\tnote{234} & 2.1 (1.1)\tnote{134} & 3.6 (0.9)\tnote{124} &
    4.0 (1.5)\tnote{123} \\
    & Rigidity & 1.5 (0.9)\tnote{234} & 1.8 (1.1)\tnote{134} & 3.3 (0.9)\tnote{124} &
    3.8 (1.4)\tnote{123} \\
    & Axial & 1.8 (1.6)\tnote{234} & 3.3 (2.1)\tnote{134} & 4.3 (2.4)\tnote{124} &
    7.4 (2.9)\tnote{123} \\
    \midrule
    Variables not & Sex (\% male) & 63 & 53\tnote{3} & 71\tnote{2} & 58 \\
    used in analysis & CISI total & 5.7 (3.2)\tnote{234} & 9.2 (3.9)\tnote{14} & 9.7
    (3.5)\tnote{14} & 14.8 (5.1)\tnote{123} \\
    & Age (34--89) & 62.7 (9.7)\tnote{34} & 64 (9.0)\tnote{4} & 65 (10.2)\tnote{14} & 70.5 (8.9)\tnote{123} \\
    & PD onset & 56 (10.6) & 55.2 (10.1) & 56.9 (11.1) & 58.2 (11.4) \\
    & PD duration & 6.7 (4.8)\tnote{234} & 8.7 (5.7)\tnote{14} & 8 (5.5)\tnote{14} &
    12.3
    (8.1)\tnote{123} \\
    \bottomrule
  \end{tabular}
  \begin{tablenotes}
    \small
    \item[1] Significant difference with cluster 1 (corrected $p < 0.05$)
    \item[2] Significant difference with cluster 2 (corrected $p < 0.05$)
    \item[3] Significant difference with cluster 3 (corrected $p < 0.05$)
    \item[4] Significant difference with cluster 4 (corrected $p < 0.05$)
  \end{tablenotes}
  \end{threeparttable}
\end{table*}

\begin{figure*}[b]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4-pub.pdf}
  \vspace{-7.5em}
  \captionsetup{justification=raggedleft,
    singlelinecheck=false
  }
  \caption{Boxplots for domains clustering.}
  \label{fig:box}
\end{figure*}

@@

$k$-means clustering performed on the 30 individual nonmotor symptoms found similar patterns to the
first clustering (Table~\ref{tab:nms30}, heatmap in Figure~\ref{fig:nms30-hm}). Means of all
symptoms were found to differ across clusters except for PD onset and tremor.

Clusters 1 ($n = 509$) and 4 ($n = 49$), the mild and severe subtypes of the previous clustering,
retained their characteristics here, while clusters 2 ($n = 97$) and 3 ($n = 249$) were
characterized differently.
Cluster 2 was a mood/apathy-dominant cluster, with a relatively higher proportion of female patients and
additionally a more severe incidence of insomnia. Cluster 3 was not characterized by any especially high
symptoms and was thus classified as a group of patients with average severity.
% Once again, CISI TOTAL...

@@

 The two clusterings were somewhat similar ($\text{ARI} = 0.46$; $\text{$v$-measure} = 0.43$).  To
 help compare the alignment of the two clusterings, we denoted the four clusters in the domains
 clustering as $D_1, D_2, D_3, D_4$, and the clusters in the symptoms clustering as $S_1, S_2, S_3,
 S_4$.
Alignment of the clusters is visualized in Figure~\ref{fig:align}. While clusters $D_1/S_1$ and
$D_4/S_4$
aligned well with each other, main differences in the clusterings occurred in the mixing of
clusters 2 and 3; for example, $D_3$ (motor-dominant) was somewhat
evenly split among $S_1$ (mild) and $S_3$ (average).

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{cluster-alignment.pdf}
  \caption{Stacked barplots indicating the distribution of opposite cluster membership for the patients of
  a given cluster. For example, the bar furthest left indicates that close to 100\% of patients in
domains cluster 1 are in symptoms cluster 1, with the rest in symptoms cluster 3.}
  \label{fig:align}
\end{figure}

@@

  statistics are reported as mean (sd).}

@@

  Lightheadedness & 0.6 (1.3)\tnote{234} & 2.4 (3.0)\tnote{14} & 2.1 (2.9)\tnote{14} & 5.2

@@

  &Fainting & 0.1 (0.7)\tnote{24} & 0.5 (1.4)\tnote{134} & 0.2 (0.6)\tnote{24} & 2.4

@@

  Drowsiness & 1.0 (1.9)\tnote{234} & 2.8 (3.3)\tnote{14} & 2.5 (3.0)\tnote{14} & 7.0

@@

  &Fatigue & 1.3 (2.0)\tnote{234} & 6.3 (4.2)\tnote{13} & 4.8 (3.9)\tnote{124} & 7.3 (3.5)\tnote{13} \\
  &Insomnia & 1.2 (2.6)\tnote{234} & 5.5 (4.8)\tnote{13} & 3.2 (3.8)\tnote{12} & 4.2

@@

  &RLS & 0.7 (1.6)\tnote{234} & 2.3 (3.5)\tnote{14} & 1.9 (3.2)\tnote{14} & 3.9

@@

  3.\ Mood/apathy &
  Loss\_interest & 0.4 (1.1)\tnote{234} & 6.1 (3.8)\tnote{134} & 1.0 (1.8)\tnote{124} & 4.6 (3.7)\tnote{123} \\
  &Loss\_activities & 0.7 (1.6)\tnote{234} & 7.1 (4.1)\tnote{134} & 1.7 (2.4)\tnote{124} &

@@

  &Anxiety & 0.9 (1.8)\tnote{234} & 6.6 (4.3)\tnote{134} & 2.0 (2.8)\tnote{124} & 4.6 (3.9)\tnote{123} \\
  &Depression & 0.8 (1.6)\tnote{234} & 7.8 (3.7)\tnote{134} & 2.3 (3.0)\tnote{124} & 4.3 (3.7)\tnote{123} \\
  &Flat\_affect & 0.4 (1.1)\tnote{234} & 4.9 (4.3)\tnote{134} & 0.9 (1.9)\tnote{124} & 2.7 (3.0)\tnote{123} \\
  &Loss\_pleasure & 0.4 (1.2)\tnote{234} & 6.8 (4.1)\tnote{134} & 1.1 (2.1)\tnote{124} & 3.4 (3.3)\tnote{123} \\

@@

  4.\ Perception/ & Hallucination & 0.2 (1.0)\tnote{234} & 0.8 (1.9)\tnote{14} & 0.6 (1.6)\tnote{14} & 4.6 (3.6)\tnote{123} \\

@@

  3.1 (3.8)\tnote{123} \\
  &Diplopia & 0.2 (0.7)\tnote{234} & 1.0 (2.5)\tnote{14} & 0.6 (1.4)\tnote{14} & 4.4

@@

  5.\ Attention/ & Loss\_concentration & 0.9 (1.8)\tnote{234} & 4.0 (3.6)\tnote{134} & 2.5 (3.1)\tnote{124} & 6.6 (3.9)\tnote{123} \\
  memory &Forget\_explicit & 0.9 (1.6)\tnote{234} & 3.8 (3.9)\tnote{134} & 2.2 (2.9)\tnote{124} & 6.7 (3.6)\tnote{123} \\
  &Forget\_implicit & 0.7 (1.6)\tnote{234} & 3.1 (3.7)\tnote{134} & 1.8 (2.6)\tnote{124} & 6.7 (4.0)\tnote{123} \\

@@

  Drooling & 0.7 (1.6)\tnote{234} & 3.0 (4.1)\tnote{14} & 3.0 (3.8)\tnote{14} & 6.1 (4.4)\tnote{123} \\
  &Swallowing & 0.4 (1.2)\tnote{234} & 1.2 (2.1)\tnote{14} & 1.7 (2.8)\tnote{14} & 4.5 (3.9)\tnote{123} \\
  &Constipation & 1.6 (2.9)\tnote{234} & 3.6 (4.3)\tnote{14} & 3.3 (4.1)\tnote{14} & 7.0 (5.1)\tnote{123} \\

@@

  Urinary\_urgency & 1.0 (1.8)\tnote{234} & 3.5 (4.1)\tnote{14} & 3.8 (3.8)\tnote{14} & 7.8 (4.0)\tnote{123} \\
  &Urinary\_frequency & 0.9 (1.9)\tnote{234} & 4.0 (4.1)\tnote{14} & 3.8 (4.0)\tnote{14} & 6.5 (4.0)\tnote{123} \\
  &Nocturia & 1.7 (2.3)\tnote{234} & 5.0 (4.3)\tnote{14} & 5.1 (4.1)\tnote{14} & 7.7 (3.9)\tnote{123} \\

@@

  Sex\_drive & 0.8 (1.9)\tnote{234} & 4.5 (4.9)\tnote{13} & 2.3 (3.6)\tnote{124} & 5.2 (5.4)\tnote{13} \\

@@

  4.6 (5.4)\tnote{13} \\

@@

  Unexplained\_pain & 0.7 (1.9)\tnote{234} & 2.5 (3.7)\tnote{14} & 3.0 (4.1)\tnote{14} & 4.2 (4.3)\tnote{123} \\

@@

  4.7 (4.8)\tnote{13} \\
  &Weight\_change & 0.9 (1.7)\tnote{234} & 2.5 (3.4)\tnote{14} & 1.9 (3.0)\tnote{14} & 3.9 (4.2)\tnote{123} \\
  &Sweating & 0.7 (1.8)\tnote{234} & 2.2 (3.5)\tnote{1} & 2.8 (4.1)\tnote{1} & 3.2

@@

  Motor symptoms &
  Tremor & 2.5 (2.3) & 2.1 (2.4) & 2.6 (2.6) & 4.1 (4.3) \\
  &Bradykinesia & 2.0 (1.2)\tnote{234} & 2.8 (1.4)\tnote{14} & 2.8 (1.4)\tnote{14} & 3.9 (1.8)\tnote{123} \\
